SyGO: Generic On-the-fly Symbolic Algorithms For Variants of GKAT

### 2025-04-16 POPL26/paper.tex
simplify semantics

Guarded NetKAT~\cite{wasserstein_GUARDEDNETKATSOUNDNESS_2023} is a variant of NetKAT that disallows non-deterministic assignments. 
For Guarded NetKAT, the possible outputs \(O\) contains acceptance \(\acc{}\) and \(\setc{σ}\), where \(σ: X ↛ V\) is a partial function that assign \(σ(x) ∈ V\) to some variable \(x ∈ X\), when \(x\) is in the domain of \(σ\).
And we can extend \(σ: X ↛ V\):
\[σ[x ↦ v](y) ≜ \begin{cases}
  v & \text{if } x = y \\  
  σ(x) & \text{otherwise}
\end{cases}\]

@@

The state in the symbolic Guarded NetKAT automata has two components: \(σ: X ↛ V\) indicating the current known assignment to variables, and \(e: \Exp\) representing the rest of the expression
\begin{mathpar}
  \inferrule
  {\\}
  {(σ, \comAssert{b})\transOut{b[σ]}{\acc{}}}
  \and  
  \inferrule
  {\\}
  {(σ, x := v)\transOut{\true}{\setc{σ[x ↦ v]}}}
  \and  
  \inferrule
  {\\}
  {(σ,\comDup)\transvia{b_σ ∣ \comDup} (σ,\comSkip)}
  \and  
  \inferrule
  {(σ, e) \transOut{c}{o} \\ o ∈ O}
  {(σ, e +_b f) \transOut{b[σ] ∧ c}{o}}
  \and  
  \inferrule
  {(σ, f) \transOut{c}{o} \\ o ∈ O}
  {(σ, e +_b f) \transOut{¬b[σ] ∧ c}{o}}
  \and
  \inferrule
  {(σ, e) \transvia{c ∣ p} (σ', e')}
  {(σ, e +_b f) \transvia{b[σ] ∧ c ∣ p} (σ', e')}
  \and  
  \inferrule
  {(σ, f) \transvia{c ∣ p} (σ', f')}
  {(σ, e +_b f) \transvia{b[σ] ∧ c ∣ p} (σ', f')}
  \and  
  \inferrule
  {(σ, e) \transOut{b}{\acc{}} \\ (σ, f) \transOut{c}{o} \\ o ∈ O}
  {(σ, e; f) \transOut{b ∧ c}{o}}
  \and  
  \inferrule
  {(σ, e) \transOut{b}{\acc{}} \\ (σ, f) \transvia{b ∣ p} (σ', f')}
  {(σ, e; f) \transvia{b ∧ c ∣ p} (σ, f')}
  \and  
  \inferrule
  {(σ, e) \transOut{b}{\setc{σ'}} \\ (σ', f) \transOut{c}{o} \\ o ∈ O}
  {(σ, e; f) \transOut{b ∧ c}{o}}
  \and  
  \inferrule
  {(σ₁, e) \transOut{b}{\setc{σ₂}} \\ (σ₂, f) \transvia{b ∣ p} (σ₃, f')}
  {(σ₁, e; f) \transvia{b ∧ c ∣ p} (σ₃, f')}
  \and  
  \inferrule
  {(σ, e) \transvia{b ∣ p} (σ', e')}
  {(σ, e;f) \transvia{b ∣ p} (σ', e';f)}
  \and  
  \inferrule
  {\\}
  {(σ, e^{(b)}) \transOut{¬b[σ]}{\acc{}}}
  \and  
  \inferrule
  {(σ₁, e) \transOut{c}{\setc{σ₂}} \\  (σ₂, e^{(b)}) \transvia{a ∣ p} (σ₃, f)}
  {(σ₁, e^{(b)}) \transvia{c ∧ a ∣ p} (σ₃, f)}
  \and  
  \inferrule
  {(σ, e) \transvia{c ∣ p} (σ', e')}
  {(σ, e^{(b)}) \transvia{b ∧ c ∣ p} (σ', e'; e^{(b)})}
\end{mathpar}

>>>>

Guarded NetKAT~\cite{wasserstein_GUARDEDNETKATSOUNDNESS_2023} is a variant of NetKAT that disallows non-deterministic assignments. 
For Guarded NetKAT, the possible outputs \(O\) are \emph{partial assignments} \(σ: X ↛ V\), which are partial functions that encodes all the \emph{known} variable assignment at this stage of the execution.
And we can extend \(σ: X ↛ V\):
\[σ[x ↦ v](y) ≜ \begin{cases}
  v & \text{if } x = y \\  
  σ(x) & \text{otherwise}
\end{cases}\]

@@

The state in the symbolic Guarded NetKAT automata has two components: \(σ: X ↛ V\) indicating the current known assignment to variables, and \(e: \Exp\) representing the rest of the expression
\begin{mathpar}
  \inferrule
  {\\}
  {(σ, \comAssert{b})\transOut{b[σ]}{σ}}
  \and  
  \inferrule
  {\\}
  {(σ, x := v)\transOut{\true}{σ[x ↦ v]}}
  \and  
  \inferrule
  {\\}
  {(σ,\comDup)\transvia{b_σ ∣ \comDup} (σ,\comSkip)}
  \and  
  \inferrule
  {(σ, e) \transOut{c}{σ'}}
  {(σ, e +_b f) \transOut{b[σ] ∧ c}{σ'}}
  \and  
  \inferrule
  {(σ, f) \transOut{c}{σ'}}
  {(σ, e +_b f) \transOut{¬b[σ] ∧ c}{σ'}}
  \and
  \inferrule
  {(σ, e) \transvia{c ∣ p} (σ', e')}
  {(σ, e +_b f) \transvia{b[σ] ∧ c ∣ p} (σ', e')}
  \and  
  \inferrule
  {(σ, f) \transvia{c ∣ p} (σ', f')}
  {(σ, e +_b f) \transvia{b[σ] ∧ c ∣ p} (σ', f')}
  \and  
  \inferrule
  {(σ₁, e) \transOut{b}{σ₂} \\ (σ₂, f) \transOut{c}{σ₃}}
  {(σ₁, e; f) \transOut{b ∧ c}{σ₃}}
  \and  
  \inferrule
  {(σ₁, e) \transOut{b}{σ₂} \\ (σ₂, f) \transvia{b ∣ p} (σ₃, f')}
  {(σ₁, e; f) \transvia{b ∧ c ∣ p} (σ₃, f')}
  \and  
  \inferrule
  {(σ, e) \transvia{b ∣ p} (σ', e')}
  {(σ, e;f) \transvia{b ∣ p} (σ', e';f)}
  \and  
  \inferrule
  {\\}
  {(σ, e^{(b)}) \transOut{¬b[σ]}{σ}}
  \and  
  \inferrule
  {(σ₁, e) \transOut{c}{σ₂} \\  (σ₂, e^{(b)}) \transOut{a}{σ₃}}
  {(σ₁, e^{(b)}) \transOut{b[σ] ∧ c ∧ a}{σ₃}}
  \and  
  \inferrule
  {(σ₁, e) \transOut{c}{σ₂} \\  (σ₂, e^{(b)}) \transvia{a ∣ p} (σ₃, f)}
  {(σ₁, e^{(b)}) \transvia{b[σ] ∧ c ∧ a ∣ p} (σ₃, f)}
  \and  
  \inferrule
  {(σ, e) \transvia{c ∣ p} (σ', e')}
  {(σ, e^{(b)}) \transvia{b[σ] ∧ c ∣ p} (σ', e'; e^{(b)})}
\end{mathpar}
Notice that when \((σ₁, e) \transOut{c}{σ₂}\), the derivative rule with the source \((σ₁, e^{(b)})\) appears to be self-referential. 
However, since there are only finitely many partial assignments \(σ: X ↛ V\), for any finite set \(X\) and \(V\); there are only finitely many cases to check.
Practically, we use the following recursive algorithm to compute \(ϵ(σ, e^{(b)})\), where \(Ψ\) denotes the set of all explored \(σ\).
\begin{align*}
  ϵ̂(σ, e^{(b)}, Ψ) & ≜ \begin{cases}
    ∅ & \text{if } σ ∈ Ψ \\  
    \begin{aligned}
      & \{(¬b[σ], σ)\} ∪ \\
      & \{(b[σ] ∧ c ∧ a, σ₂) ∣ (a, σ₂) ∈ ϵ̂(σ₁, e^{(b)}, Ψ ∪ \{σ\})\}
    \end{aligned}
     & \text{else, if } (σ, e)\transOut{c}{σ₁}\\
  \end{cases}\\
  ϵ(σ, e^{(b)}) & ≜ ϵ̂(σ, e^{(b)}, ∅)
\end{align*}
